This page and the next page should be removed

### 2021-01-16 analysis/manuscript.Rmd
before moving the preproc step into an r chunk

```{r exp1-preparation}
# Define function for reading and formatting behavioral data
read_behav <- function(txt_fname) {
  # Read log file
  dat <- suppressWarnings(read_tsv(txt_fname, col_types = cols()))
  # Rename columns for participant and item IDs
  dat %<>% rename(participant = VPNummer, item = StimID) %>%
    # Add a column for the part of the experiment
    mutate(part = case_when(
      Wdh %in% c(211, 231) ~ "I",
      Wdh %in% c(212, 232) ~ "II",
      Wdh %in% c(213, 233) ~ "III",
      Wdh == 234 ~ "IV"
    ) %>% as_factor()) %>%
    # Remove filler stimuli (i.e.,well-known objects)
    filter(bek_unbek != "bekannt")
  # Assign stimuli to conditions based on keywords and button presses
  excl_known <- dat %>% filter(part == "I" & Tastencode %in% c(201, 251)) %>% pull(item)
  excl_informed <- dat %>% filter(part == "II" & grepl("richtig", Bed) & !Tastencode %in% c(201, 202, 251, 252)) %>% pull(item)
  excl_naive <- dat %>% filter(part == "II" & grepl("falsch", Bed) & !Tastencode %in% c(203, 204, 253, 254)) %>% pull(item)
  informed <- dat %>% filter(part == "II" & grepl("richtig", Bed) & Tastencode %in% c(201, 202, 251, 252)) %>% pull(item)
  naive <- dat %>% filter(part == "II" & grepl("falsch", Bed) & Tastencode %in% c(203, 204, 253, 254)) %>% pull(item)
  # Create a new condition for the experimental condition
  dat %<>% mutate(condition = case_when(
    item %in% excl_known ~ "Excl_known",
    item %in% excl_informed ~ "Excl_informed",
    item %in% excl_naive ~ "Excl_naive",
    item %in% informed ~ "Informed",
    item %in% naive ~ "Naive"
  ) %>% factor(levels = c("Informed", "Naive", "Excl_informed", "Excl_naive", "Excl_known")))
  # In Experiment 3, we need an additional column for position (upright vs. inverted)
  if ("richtig_inv" %in% dat$Bed) {
    dat %<>% mutate(position = as_factor(ifelse(grepl("inv", Bed), "Inverted", "Upright"))) %>%
      # Return only relevant columns
      select(part, position, condition, participant, item, RT)
  } else {
    # Return only relevant columns
    dat %<>% select(part, condition, participant, item, RT)
  }
  return(dat)
}

@@

# Check average number of stimuli per condition
(stims_exp1 <- map(data_exp1, function(x) {table(x$condition) / 3}) %>% bind_rows() %>% colMeans())
```

```{python exp1-preprocessing}
# Import libraries
import mne
import glob
import os

montage = mne.channels.make_standard_montage("easycap-M1")
montage._get_ch_pos()

# Set parameters for EEG preprocessing
preproc_params = dict(
    n_components=15, method="fastica",         # ICA parameters
    l_freq=0.1, h_freq=30,                     # Filter edges
    event_id={"match": 221, "mismatch": 222},  # EEG triggers
    tmin=-0.5, tmax=1.498,                     # Length of epochs
    baseline=(-0.2, 0),                        # Baseline correction
    reject=dict(eeg=200)                       # Rejection threshold
)

# # Debug
# locals().update(preproc_params)
# vhdr_fname = 'data/exp1/EEG/Vp0001.vhdr'

# Define function for preprocessing EEG data
def preproc(vhdr_fname, metadata, n_components, method, l_freq, h_freq, event_id, tmin, tmax, baseline, reject):
    # Read EEG data in microvolts
    raw = mne.io.read_raw_brainvision(vhdr_fname, scale=1e6, preload=True)
    # Create virtual EOG channels
    raw = mne.set_bipolar_reference(raw, "Auge_u", "Fp1", ch_name="VEOG", drop_refs=False)
    raw = mne.set_bipolar_reference(raw, "F9", "F10", ch_name="HEOG", drop_refs=False)
    raw.set_channel_types(mapping={"VEOG": "eog", "HEOG": "eog"})
    # Add EasyCap electrode layout, removing any excessive channels
    montage = mne.channels.make_standard_montage("easycap-M1")
    raw.drop_channels(list(set(raw.ch_names) - set(montage.ch_names) - set(["VEOG", "HEOG"])))
    raw.set_montage(montage=montage)
    # # Interpolate bad channels
    # if vhdr_fname=='data/exp3/EEG/VP05.vhdr':
    #   raw.info['bads'].extend(['AF8', 'TP9'])
    #   raw = raw.interpolate_bads()
    # Re-reference to common average
    raw, _ = mne.set_eeg_reference(raw, "average")
    # Run ICA on a copy of the data
    filt_raw = raw.copy()
    filt_raw.load_data().filter(l_freq=1, h_freq=None)
    ica = mne.preprocessing.ICA(n_components=n_components, random_state=12345, method=method)
    ica.fit(filt_raw)
    # Remove bad components based on correlations with EOG
    eog_indices, eog_scores = ica.find_bads_eog(raw)
    ica.exclude = eog_indices
    raw = ica.apply(raw)
    # Apply band-pass filter
    raw = raw.filter(l_freq=l_freq, h_freq=h_freq)
    # Epoching including baseline correction
    events, _ = mne.events_from_annotations(raw, verbose=False)
    epochs = mne.Epochs(raw, events=events, event_id=event_id, tmin=tmin, tmax=tmax, baseline=baseline, preload=True)
    # Add behavioral data
    epochs.metadata = metadata
    # Reject bad epochs
    epochs = epochs.drop_bad(reject=reject)
    return epochs

# List raw EEG filenames
fnames_exp1 = sorted(glob.glob("data/exp1/EEG/*.vhdr"))

@@

# Check if preprocessing was done already (delete the file to re-run)
if os.path.exists("analysis/export/exp1-epo.fif"):
    # Read preprocessed data from file
    epochs_exp1 = mne.read_epochs("analysis/export/exp1-epo.fif", preload=True)
else:
    # Preprocess EEG data for Experiment 1
    epochs_exp1 = [
        preproc(vhdr_fname=fname, metadata=meta, **preproc_params) for fname, meta in zip(fnames_exp1, metadata_exp1)
    ]
    # Combine epochs into a single data set
    epochs_exp1 = mne.concatenate_epochs(epochs_exp1)
    # Backup epochs to the export folder
    epochs_exp1.save("analysis/export/exp1-epo.fif")

# Define function to compute grand averages per condition
def compute_evokeds(epochs):
    evokeds = dict(); evokeds_dat = dict()
    # Each part becomes a dictionary of conditions
    for pt in ["I", "II", "III"]:
        evokeds[pt] = dict(); evokeds_dat[pt] = dict()
        # Each condition becomes a list of participants
        for cn in ["Informed", "Naive"]:
            evokeds[pt][cn] = list(); evokeds_dat[pt][cn] = list()
            # For every participant we average trials separately for all parts and conditions
            for vp in epochs.metadata["participant"].unique():
                query = 'participant == "' + vp + '" and part == "' + pt + '" and condition == "' + cn + '"'
                evokeds[pt][cn].append(epochs[query].average())
            # Compute grand averages across participants for this part and condition
            evokeds[pt][cn] = mne.grand_average(evokeds[pt][cn])
            # Export only the actual data for R
            evokeds_dat[pt][cn] = evokeds[pt][cn].data
    return (evokeds, evokeds_dat)

# Compute grand averages for Experiment 1
evokeds_exp1, evokeds_dat_exp1 = compute_evokeds(epochs=epochs_exp1)
```

>>>>

```{r exp1-preparation, eval=rerun_preproc}
# Define function for reading and formatting the behavioral data
read_behav <- function(txt_fname) {
  # Read the log file
  dat <- suppressWarnings(read_tsv(txt_fname, col_types = cols()))
  # Rename columns for participant and item IDs
  dat %<>% rename(participant = VPNummer, item = StimID) %>%
    # Add a column for the part of the experiment
    mutate(part = case_when(
      Wdh %in% c(211, 231) ~ "I",
      Wdh %in% c(212, 232) ~ "II",
      Wdh %in% c(213, 233) ~ "III",
      Wdh == 234 ~ "IV"
    ) %>% as_factor()) %>%
    # Remove filler stimuli (i.e.,well-known objects)
    filter(bek_unbek != "bekannt")
  # Assign stimuli to conditions based on keywords and button presses
  excl_known <- dat %>% filter(part == "I" & Tastencode %in% c(201, 251)) %>% pull(item)
  excl_informed <- dat %>% filter(part == "II" & grepl("richtig", Bed) & !Tastencode %in% c(201, 202, 251, 252)) %>% pull(item)
  excl_naive <- dat %>% filter(part == "II" & grepl("falsch", Bed) & !Tastencode %in% c(203, 204, 253, 254)) %>% pull(item)
  informed <- dat %>% filter(part == "II" & grepl("richtig", Bed) & Tastencode %in% c(201, 202, 251, 252)) %>% pull(item)
  naive <- dat %>% filter(part == "II" & grepl("falsch", Bed) & Tastencode %in% c(203, 204, 253, 254)) %>% pull(item)
  # Create a new condition for the experimental condition
  dat %<>% mutate(condition = case_when(
    item %in% excl_known ~ "Excl_known",
    item %in% excl_informed ~ "Excl_informed",
    item %in% excl_naive ~ "Excl_naive",
    item %in% informed ~ "Informed",
    item %in% naive ~ "Naive"
  ) %>% factor(levels = c("Informed", "Naive", "Excl_informed", "Excl_naive", "Excl_known")))
  # In Experiment 3, we need an additional column for position (upright vs. inverted)
  if ("richtig_inv" %in% dat$Bed) {
    dat %<>% mutate(position = as_factor(ifelse(grepl("inv", Bed), "Inverted", "Upright"))) %>%
      # Return only relevant columns
      select(part, position, condition, participant, item, RT)
  } else {
    # Return only relevant columns
    dat %<>% select(part, condition, participant, item, RT)
  }
  return(dat)
}

@@

# Check average number of stimuli per condition and save to the export folder
map(data_exp1, function(x) table(x$condition) / 3) %>%
  bind_rows() %>%
  colMeans() %>%
  saveRDS("analysis/export/exp1-stims.RDS")
```

```{python exp1-preprocessing, eval=rerun_preproc}
# Import libraries
import mne
import glob
import os

# Set parameters for EEG preprocessing
preproc_params = dict(
    n_components=15, method='fastica',         # ICA parameters
    l_freq=0.1, h_freq=30,                     # Filter edges
    event_id={'match': 221, 'mismatch': 222},  # EEG triggers
    tmin=-0.5, tmax=1.498,                     # Length of epochs
    baseline=(-0.2, 0),                        # Baseline correction
    reject=dict(eeg=200)                       # Rejection threshold
)

# # For debugging
# locals().update(preproc_params)
# vhdr_fname = 'data/exp1/EEG/Vp0001.vhdr'

# Define function for preprocessing the EEG data
def preproc(vhdr_fname, metadata, n_components, method, l_freq, h_freq,
            event_id, tmin, tmax, baseline, reject):
    # Read EEG data in microvolts
    raw = mne.io.read_raw_brainvision(vhdr_fname, scale=1e6, preload=True)
    # Create virtual EOG channels
    raw = mne.set_bipolar_reference(raw, 'Auge_u', 'Fp1', ch_name='VEOG', drop_refs=False)
    raw = mne.set_bipolar_reference(raw, 'F9', 'F10', ch_name='HEOG', drop_refs=False)
    raw.set_channel_types(mapping={'VEOG': 'eog', 'HEOG': 'eog'})
    # Add EasyCap electrode layout, removing any excessive channels
    montage = mne.channels.make_standard_montage('easycap-M1')
    raw.drop_channels(list(set(raw.ch_names) - set(montage.ch_names) - set(['VEOG', 'HEOG'])))
    raw.set_montage(montage=montage)
    # Re-reference to common average
    raw, _ = mne.set_eeg_reference(raw, 'average')
    # Run ICA on a copy of the data
    filt_raw = raw.copy()
    filt_raw.load_data().filter(l_freq=1, h_freq=None)
    ica = mne.preprocessing.ICA(n_components=n_components, random_state=12345, method=method)
    ica.fit(filt_raw)
    # Remove bad components based on correlations with EOG
    eog_indices, eog_scores = ica.find_bads_eog(raw)
    ica.exclude = eog_indices
    raw = ica.apply(raw)
    # Apply band-pass filter
    raw = raw.filter(l_freq=l_freq, h_freq=h_freq)
    # Epoching including baseline correction
    events, _ = mne.events_from_annotations(raw, verbose=False)
    epochs = mne.Epochs(raw, events=events, event_id=event_id, tmin=tmin,
                        tmax=tmax, baseline=baseline, preload=True)
    # Add behavioral data
    epochs.metadata = metadata
    # Reject bad epochs
    epochs = epochs.drop_bad(reject=reject)
    return epochs

# List raw EEG filenames
fnames_exp1 = sorted(glob.glob('data/exp1/EEG/*.vhdr'))

@@

# Apply the function to preprocess the EEG data for Experiment 1
epochs_exp1 = [preproc(vhdr_fname=fname, metadata=meta, **preproc_params)
               for fname, meta in zip(fnames_exp1, metadata_exp1)]

# Combine epochs into a single data set
epochs_exp1 = mne.concatenate_epochs(epochs_exp1)

# Backup epochs to the export folder
epochs_exp1.save('analysis/export/exp1-epo.fif')

# Backup the channel names separately

# Check if preprocessing was done already (delete the file to re-run)
if os.path.exists('analysis/export/exp1-epo.fif'):
    # Read preprocessed data from file
    epochs_exp1 = mne.read_epochs('analysis/export/exp1-epo.fif', preload=True)
else:
    # Preprocess EEG data for Experiment 1
    epochs_exp1 = [preproc(vhdr_fname=fname, metadata=meta, **preproc_params)
                   for fname, meta in zip(fnames_exp1, metadata_exp1)]
    # Combine epochs into a single data set
    epochs_exp1 = mne.concatenate_epochs(epochs_exp1)
    # Backup epochs to the export folder
    epochs_exp1.save('analysis/export/exp1-epo.fif')

# Define function to compute grand averages per condition
def compute_evokeds(epochs):
    evokeds = dict(); evokeds_dat = dict()
    # Each part becomes a dictionary of conditions
    for pt in ['I', 'II', 'III']:
        evokeds[pt] = dict(); evokeds_dat[pt] = dict()
        # Each condition becomes a list of participants
        for cn in ['Informed', 'Naive']:
            evokeds[pt][cn] = list(); evokeds_dat[pt][cn] = list()
            # For every participant we average trials separately for all parts and conditions
            for vp in epochs.metadata['participant'].unique():
                query = 'participant == "' + vp + '" and part == "' + pt + '" and condition == "' + cn + '"'
                evokeds[pt][cn].append(epochs[query].average())
            # Compute grand averages across participants for this part and condition
            evokeds[pt][cn] = mne.grand_average(evokeds[pt][cn])
            # Export only the actual data for R
            evokeds_dat[pt][cn] = evokeds[pt][cn].data
    return (evokeds, evokeds_dat)

# Compute grand averages for Experiment 1
evokeds_exp1, evokeds_dat_exp1 = compute_evokeds(epochs=epochs_exp1)

import pickle
with open('analysis/export/exp1-avgs.pickle', 'wb') as handle:
    pickle.dump(evokeds_dat_exp1, handle, protocol=pickle.HIGHEST_PROTOCOL)
```

```{r backup}

test <- py$preproc(
  vhdr_fname = "data/exp1/EEG/Vp0001.vhdr",
  metadata = data_exp1[[1]],
  n_components = as.integer(15),
  method = "fastica",
  l_freq = 0.1,
  h_freq = 30,
  event_id = list(match = as.integer(221), mismatch = as.integer(222)),
  tmin = -0.5,
  tmax = 1.498,
  baseline = tuple(-0.2, 0),
  reject = list(eeg = 200))

test$ch_names
```
